Evolution of precopulatory and postcopulatory inbreeding behaviour

### 2016-09-06 ms_pre_post.Rmd
Add second to last paragraph of introductionand associated references

<!--- TODO: GENERAL PARAGRAPH INTRODUCING THE SCOPE AND KEY QUESTION --->
Reproductive decisions expressed among interacting individuals directly affect individual fitnesses and the genotypes and viabilities of resulting offspring, and thereby profoundly affect evolutionary dynamics through the differential inheritence of alleles over generations [@Pizzari2012; @Wright2013; @Wild2014; @Liao2015; @Reid2016]. For example, how often a focal female mates is a reproductive decision that can directly and indirectly affect her fitness, and potentially alter the distribution of genotypes and patterns of inheritence within a population [e.g., @Cornell2007; @Diaz2010; @Okada2015; @Roberts2015a; @Worthington2016]. Her mating frequency and expression of mate choice might be affected by the genetic quality or complementarity of available potential mates, and her choosiness might additionally be expressed through post-copulatory mechanisms [@Parker1983; @Parker2006; @Kuijper2012]. Female choice might therefore be realised by evolution of multiple, potentially interdependent, behavioural and physiological phenotypes operating in tandem. However, no general modelling framework predicts the conditions under which polyandry, pre-copulatory choice, and post-copulatory choice evolve together, meaning that there are no formal hypotheses for predicting the relevance of these reproductive decisions across different biological systems.

<!--- TODO: PARAGRAGH ON INBREEDING, PRE-COP, AND POST-COP AVOIDANCE --->
Inbreeding, defined as breeding between relatives, directly affects inheritence of identical-by-descent alleles and the viability of inbred offspring, meaning that the decision to prefer, avoid, or tolerate (i.e., mate randomly with respect to relatedness) inbreeding is predicted to have profound effects on individual fitness [@Parker2006; @Kokko2006; @Szulkin2012; @Duthie2015a]. Viability of inbred offspring typically decreases with increasing relatedness between parents [termed 'inbreeding depression'; @Keller2002; @Charlesworth2009]. If inbreeding depression is sufficiently weak, selection for female inbreeding preference is predicted because inbreeding increases the reproductive success of a focal female's male relative, thereby increasing the number of identical-by-descent copies present in her offspring [@Parker1979; @Parker2006; @Kokko2006; @Duthie2015a]. However, if inbreeding depression is moderate to strong, selection for female inbreeding avoidance is predicted [@Duthie]. Female inbreeding strategy can be enacted before or after copulation occurs [e.g., @Tuni2013; @Firman2015; @Daniel2015; @Fischer2015]. However, models of inbreeding do not consider pre-copulatory inbreeding strategy (hereafter 'inbreeding mating strategy') and post-copulatory inbreeding strategy (hereafter 'inbreeding fertilisation strategy') jointly, meaning that it is unclear when selection for either or both strategies is predicted in empirical systems.

@@

<!--- TODO: PARAGRAGH ON EVOLUTION OF POLYANDRY --->
Critically, evolution of post-copulatory inbreeding avoidance is precluded if females only mate with one male, meaning that evolution of polyandry is intrinsically tied to female inbreeding fertilisation strategy. Although polyandry is commonplace in wild populations, the regularity with which it is observed in the wild remains theoretically puzzling because engaging in polyandry does not necessarily increase the number of offspring that a female produces, meaning that the fitness benefits associated with polyandry are not always clear [@Parker2013; @Taylor2014]. It has been hypothesised that females can acheive a net fitness benefit if polyandry provides them with additional opportunities to adjust their inbreeding through copulations with additional males [e.g., @Tregenza2002; @Akcay2007; @Lehtonen2015]. @Duthie showed that selection for polyandry to facilitate inbreeding avoidance is sometimes predicted, but only under highly constrained conditions in which very few males are available during a female's initial mate choice but many males are available during a female's additional mate choice(s), or when expression of polyandry is conditionally expressed based on the degree of inbreeding in a female's initially chosen mate. However, @Duthie assumed that polyandrous females had no control over the assignment of paternity to their offspring. If after engaging in polyandry, females could bias paternity toward non-relatives, then selection for polyandry to avoid inbreeding might be more widely predicted [@Zeh1997]. However, no models quantify the degree to which expression of biased paternity resulting from evolving inbreeding fertilisation strategy can feedback to drive evolution of polyandry in populations of potentially inbreeding females and males with biologically realistic distributions of relatedness.

@@

@@

<!--- Change mating cost to an increased risk of mating failure ---> 
Values of phenotypic traits for tendency to engage in polyandry ($P_{p}$), mating ($M_{p}$), and fertilistation ($F_{p}$) each have costs, which independently increase the probability of a focal female's reproductive failure. These costs are all expressed prior to mating. In the case of tendency to engage in polyandry, the cost of $P_{p}$ ($c_{P}$) is interpreted biologically as a cost associated with mate searching or courtship [e.g., @Rowe1988; @Rowe1994; @Ronkainen1994; @Koga1998; @Lasley-Rasher2012]. For inbreeding mating strategy, the cost of $M_{p}$ ($c_{M}$) is interpreted biologically as a cost associated with risk of mating failure due to females being overly choosy [e.g., @Kokko2013], and for inbreeding fertilisation strategy, the cost of $F_{p}$ ($c_{F}$) is interpreted as an energetic cost associated with the physiological or biochemical mechanisms that need to be developed to bias fertilsation for inbreeding avoidance or inbreeding preference [e.g., @Gasparini2011]. Each cost independently increases the risk of female pre-mating mortality and, because generations are non-overlapping, total reproductive failure as a linear function of its associated trait. The probability that a focal female dies before mating due to the cost of polyandry is therefore $P_{p} \times c_{P}$. Similarly, the probability of pre-mating female mortality due to the cost of mating strategy and fertilisation strategy is $|M_{p}| \times c_{M}$ and $|F_{p}| \times c_{F}$, repectively, where $|M_{p}|$ is the absolute value of $M_{p}$ and $|F_{p}|$ is the absolute value of $F_{p}$. Absolute values are used for applying costs to inbreeding strategies because both negative and positive $M_{p}$ and $F_{p}$ bias breeding interactions with respect to kinship, while only positive values of $P_{p}$ affect a female's tendency for polyandry; zero and negative $P_{p}$ values always translate to monandrous phenotypes (see below).

@@

To test our hypothesis that selection for evolution of polyandry is driven by female's ability to bias fertilisation toward inbreeding avoidance through post-copulatory mechanisms, we compare simulations in which all three phenotypes including tendency for polyandry ($P_{p}$), inbreeding mating strategy ($M_{p}$), and inbreeding fertilisation strategy ($F_{p}$) evolve, with an identical set of simulations in which $F_{p}$ cannot evolve. In simulations without evolving $F_{p}$, we therefore include all 16 combinations of $c_{P}=\{0,0.02\}$, $c_{M}=\{0,0.02\}$, $c_{F}=\{0,0.02\}$, and $\beta = 3$ replicated 40 times each. To stop $F_{p}$ from evolving, we sever the connection from $F_{a}$ to $F_{p}$ such that all $F_{g}$ genotypes resulting from $F_{a}$ cause random fertilisation with respect to kinship. Hence, while mutation and drift can still occur for $F_{a}$ alleles, these alleles have no phenotypic effect and polyandry cannot cause polyandrous females to benefit through an increased ability to bias fertilisation toward inbreeding avoidance (under such conditions, there can also never be a cost of $F_{p}$, meaning that $c_{F}=0$). By comparing evolution of mean $P_{p}$ values over generations when $F_{p}$ also evolves versus when $F_{p}=0$ (random fertilisation), we test whether or not evolution of inbreeding fertilisation strategy increases selection for evolution of polyandry.

>>>>

<!--- TODO: GENERAL PARAGRAPH INTRODUCING THE SCOPE AND KEY QUESTION --->
Reproductive decisions expressed among interacting individuals directly affect individual fitnesses and the genotypes and viabilities of resulting offspring, and thereby profoundly affect evolutionary dynamics through the differential inheritence of alleles over generations [@Pizzari2012; @Wright2013; @Wild2014; @Liao2015; @Reid2016]. For example, how often a focal female mates is a reproductive decision that can directly and indirectly affect her fitness, and potentially alter the distribution of genotypes and patterns of inheritence within a population [e.g., @Cornell2007; @Diaz2010; @Okada2015; @Roberts2015a; @Worthington2016]. Her mating frequency and expression of mate choice might be affected by the genetic quality or complementarity of available potential mates, and her choosiness might additionally be expressed after mating has taken place through post-copulatory mechanisms [@Parker1983; @Parker2006; @Kuijper2012]. Female choice might therefore be realised by evolution of multiple, potentially interdependent, behavioural and physiological phenotypes operating in tandem. However, no general modelling framework predicts the conditions under which polyandry, pre-copulatory choice, and post-copulatory choice evolve together, meaning that there are no formal hypotheses for predicting the relevance of these reproductive decisions across different biological systems.

<!--- TODO: PARAGRAGH ON INBREEDING, PRE-COP, AND POST-COP AVOIDANCE --->
Inbreeding, defined as breeding between relatives, directly affects inheritence of identical-by-descent alleles and the viability of inbred offspring, meaning that the decision to prefer, avoid, or tolerate (i.e., mate randomly with respect to relatedness) inbreeding is predicted to have profound effects on individual fitness [@Parker2006; @Kokko2006; @Szulkin2012; @Duthie2015a]. Viability of inbred offspring typically decreases with increasing relatedness between parents [termed 'inbreeding depression'; @Keller2002; @Charlesworth2009]. If inbreeding depression is sufficiently weak, selection for female inbreeding preference is predicted because inbreeding increases the reproductive success of a focal female's male relative, thereby increasing the number of identical-by-descent copies present in her offspring [@Parker1979; @Parker2006; @Kokko2006; @Duthie2015a]. However, if inbreeding depression is moderate to strong, selection for female inbreeding avoidance is predicted [@Duthie; Duthie and Reid *in press*]. Female inbreeding strategy can be enacted before or after copulation occurs [e.g., @Tuni2013; @Firman2015; @Daniel2015; @Fischer2015]. However, models of inbreeding do not consider pre-copulatory inbreeding strategy (hereafter 'inbreeding mating strategy') and post-copulatory inbreeding strategy (hereafter 'inbreeding fertilisation strategy') jointly, meaning that it is unclear when selection for either or both strategies is predicted in empirical systems.

@@

<!--- TODO: PARAGRAGH ON EVOLUTION OF POLYANDRY --->
Critically, evolution of post-copulatory inbreeding strategy is precluded if females only mate with one male, meaning that evolution of polyandry is intrinsically tied to female inbreeding fertilisation strategy. Although polyandry is commonplace in wild populations, the regularity with which it is observed in the wild remains theoretically puzzling because engaging in polyandry does not necessarily increase the number of offspring that a female produces, meaning that the fitness benefits associated with polyandry are not always clear [@Parker2013; @Taylor2014]. It has been hypothesised that females can acheive a net fitness benefit if polyandry provides them with additional opportunities to adjust their inbreeding through copulations with additional males [e.g., @Tregenza2002; @Akcay2007; @Lehtonen2015]. @Duthie showed that selection for polyandry to facilitate inbreeding avoidance is sometimes predicted, but only under highly constrained conditions in which very few males are available during a female's initial mate choice but many males are available during a female's additional mate choice(s), or when expression of polyandry is conditionally expressed based on the degree of inbreeding in a female's initially chosen mate. However, @Duthie assumed that polyandrous females had no control over the assignment of paternity to their offspring. If after engaging in polyandry, females could bias paternity toward non-relatives, then selection for polyandry to avoid inbreeding might be more widely predicted [@Zeh1997]. However, no models quantify the degree to which expression of biased paternity resulting from evolving inbreeding fertilisation strategy can feedback to drive evolution of polyandry in populations of potentially inbreeding females and males with biologically realistic distributions of relatedness.

@@

<!--- TODO: PARAGRAGH ON ORDER OF INTERACTIONS --->
Selection for inbreeding mating strategy versus inbreeding fertilisation strategy and associated polyandry might depend evolutionary constraints, or on existing phentypes in the population. For example, if phenotypes biasing fertilisation to avoid inbreeding are precluded by evolutionary constraints [@Bradshaw1991; @Futuyma2010], then inbreeding avoidance might only be possible before copulation occurs, perhaps strengthening selection for inbreeding mating strategy and weakening selection for polyandry. Alternatively, evolution of female inbreeding mating strategy might be precluded due to severely limited kin recognition or the inability to resist male coercion given sexual conflict [Arnqvist2005a; Parker2006; @Penn2010]. In such conditions, females might only be able to express choice through inbreeding fertilisation strategy, perhaps strengthening selection for inbreeding avoidance through polyandry and post-copulatory mechanisms. Selection for inbreeding mating or fertilisation strategy might also be weak or negligible if the strategy is redundant due to the presence of an already established adaptive inbreeding strategy (e.g., if females already successfully avoid inbreeding during mate encounters, then they might be unable to increase their fitness further by biasing fertilisation after mating has occurred). It is therefore necessary to consider evolution of inbreeding mating strategy and inbreeding fertilisation strategy when each evolves in a population where its complementary inbreeding strategy is evolutionarily constrained or already fixed as an adaptation.

@@

<!--- Change mating cost to an increased risk of mating failure ---> 
Values of phenotypic traits for tendency to engage in polyandry ($P_{p}$), mating ($M_{p}$), and fertilistation ($F_{p}$) each have costs, which independently increase the probability of a focal female's reproductive failure. These costs are all expressed prior to mating. In the case of tendency to engage in polyandry, the cost of $P_{p}$ ($c_{P}$) is interpreted biologically as a cost associated with mate searching or courtship [e.g., @Rowe1988; @Rowe1994; @Ronkainen1994; @Koga1998; @Lasley-Rasher2012]. For inbreeding mating strategy, the cost of $M_{p}$ ($c_{M}$) is interpreted biologically as a cost associated with risk of mating failure due to females being overly choosy [e.g., @Kokko2013], and for inbreeding fertilisation strategy, the cost of $F_{p}$ ($c_{F}$) is interpreted as an energetic cost associated with the physiological or biochemical mechanisms that need to be developed to bias fertilisation for inbreeding avoidance or inbreeding preference [e.g., @Gasparini2011]. Each cost independently increases the risk of female pre-mating mortality and, because generations are non-overlapping, total reproductive failure as a linear function of its associated trait. The probability that a focal female dies before mating due to the cost of polyandry is therefore $P_{p} \times c_{P}$. Similarly, the probability of pre-mating female mortality due to the cost of mating strategy and fertilisation strategy is $|M_{p}| \times c_{M}$ and $|F_{p}| \times c_{F}$, repectively, where $|M_{p}|$ is the absolute value of $M_{p}$ and $|F_{p}|$ is the absolute value of $F_{p}$. Absolute values are used for applying costs to inbreeding strategies because both negative and positive $M_{p}$ and $F_{p}$ bias breeding interactions with respect to kinship, while only positive values of $P_{p}$ affect a female's tendency for polyandry; zero and negative $P_{p}$ values always translate to monandrous phenotypes (see below).

@@

To test our hypothesis that selection for evolution of polyandry is driven by female's ability to bias fertilisation toward inbreeding avoidance through post-copulatory mechanisms, we compare simulations in which all three phenotypes including tendency for polyandry ($P_{p}$), inbreeding mating strategy ($M_{p}$), and inbreeding fertilisation strategy ($F_{p}$) evolve, with an identical set of simulations in which $F_{p}$ cannot evolve. In simulations without evolving $F_{p}$, we therefore include all 8 combinations of $c_{P}=\{0,0.02\}$, $c_{M}=\{0,0.02\}$, $c_{F}=\{0,0.02\}$, and $\beta = 3$ replicated 40 times each. To stop $F_{p}$ from evolving, we sever the connection from $F_{a}$ to $F_{p}$ such that all $F_{g}$ genotypes resulting from $F_{a}$ cause random fertilisation with respect to kinship. Hence, while mutation and drift can still occur for $F_{a}$ alleles, these alleles have no phenotypic effect and polyandry cannot cause polyandrous females to benefit through an increased ability to bias fertilisation toward inbreeding avoidance (under such conditions, there can also never be a cost of $F_{p}$, meaning that $c_{F}=0$). By comparing evolution of mean $P_{p}$ values over generations when $F_{p}$ also evolves versus when $F_{p}=0$ (random fertilisation), we test whether or not evolution of inbreeding fertilisation strategy increases selection for evolution of polyandry.
